---
title: "Tidy Tuesday "
output: 
  html_document:
    toc: true
    code_folding: hide
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(plotly)
library(ggrepel)
sports <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-29/sports.csv')

```


# Introduction + Question


Q: do gender disparities in expenditure per participant differ by sport. Particular examine this disparity int he context of gender participation ratios by sport. 

* for now do longitudinal and maybe consider school characterization (d1 vs d3...)

# Data

1. claculat expenditure per paritipant by gender
2. calculate gender participation (pct of particiaptns that are women)

we capture disparity in expenditure as a ratio: (expenditure per male participant)/(expenditure per female participant). A value of 1 is parity. A value greater thatn 1 indicates men have high expenditure per participant. A value less than 1 indicates female have higher.

```{r}
## Subset to columns we are intersted
df1 = sports %>% 
  select(id = unitid,classification_name,year,
         sport = sports,
         partic_men, partic_women,
         exp_men, exp_women)

## Filter for 2019 and coed sports
df2 = df1 %>%
  filter(year == max(year), ##for most recent year
         !is.na(partic_men)&!is.na(partic_women) ## coed sports
         ) %>% 
  select(-year)
 

## Claculate metrics
df3 = df2 %>% 
  mutate(
    exp_per_partic_men = exp_men/partic_men,
    exp_per_partic_women = exp_women/partic_women,
    ratio_exp_per_partic = exp_per_partic_men/exp_per_partic_women,
    pct_partic_women = partic_women/(partic_women + partic_men)*100
  ) 

## Analytics dataset
df_school = df3 %>% 
  select(id, sport,
         ratio_exp_per_partic,
         pct_partic_women
         )

df_sport = df3 %>% 
  group_by(sport) %>% 
  summarize_at(vars(c(exp_per_partic_men,exp_per_partic_women,
                      ratio_exp_per_partic,pct_partic_women
                      )),~mean(.x)) %>% 
  ungroup()
```



# Visualization

Lets first make mini plots. First relationship between disparitiy in expenditure per particupant and pct of women participating. 

```{r}

df_sport %>% 
  ggplot(aes(y = ratio_exp_per_partic,
            x = pct_partic_women,
            label = sport))+
  geom_point()+
  geom_text_repel()+
  geom_hline(yintercept = 0.7, lty=2)+
  geom_hline(yintercept = 1.3, lty = 2)+
  geom_hline(yintercept = 1)+
  scale_y_continuous(limits = c(0,1.5))+
  scale_x_continuous(limits = c(0,100))+
  labs(
    x = "% of participants that are women",
    y = "Expenditure disparity by gender",
    subtitle = "Disparity between expenditure per participant by gender",
    title = "% female participants vs expenditure disparity"
  )+
  theme_classic()

```

Lets look at differences

```{r}
df_sport %>% 
  mutate(diff = exp_per_partic_men - exp_per_partic_women) %>% 
  ggplot(aes(y = ratio_exp_per_partic,
            x = pct_partic_women,
            label = sport))+
  geom_point()+
  geom_text_repel()+
  geom_hline(yintercept = 0.7, lty=2)+
  geom_hline(yintercept = 1.3, lty = 2)+
  geom_hline(yintercept = 1)+
  scale_y_continuous(limits = c(0,1.5))+
  scale_x_continuous(limits = c(0,100))+
  labs(
    x = "% of participants that are women",
    y = "Expenditure disparity by gender",
    subtitle = "Disparity between expenditure per participant by gender",
    title = "% female participants vs expenditure disparity"
  )+
  theme_classic()
```

