---
title: "Tidy Tuesday "
output: 
  html_document:
    toc: true
    code_folding: hide
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
sports <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-29/sports.csv')

```


# Introduction + Question


Q: do gender disparities in expenditure per participant differ by sport. Particular examine this disparity int he context of gender participation ratios by sport. 

* for now do longitudinal and maybe consider school characterization (d1 vs d3...)

# Data

1. claculat expenditure per paritipant by gender
2. calculate gender participation (pct of particiaptns that are women)

```{r}
## Subset to columns we are intersted
df1 = sports %>% 
  select(id = unitid,classification_name,year,
         sport = sports,
         partic_men, partic_women,
         exp_men, exp_women)

## Filter for 2019 and coed sports
df2 = df1 %>%
  filter(year == max(year), ##for most recent year
         !is.na(partic_men)&!is.na(partic_women) ## coed sports
         ) %>% 
  select(-year)
 

## Claculate metrics
df3 = df2 %>% 
  mutate(
    exp_per_partic_men = exp_men/partic_men,
    exp_per_partic_women = exp_women/partic_women,
    pct_partic_women = partic_women/(partic_women + partic_men)
  ) 

## Analytics dataset
dfa = df3 %>% 
  select(id, sport,
         exp_per_partic_men,
         exp_per_partic_women,
         pct_partic_women
         )
```



# Visualization

Lets first make mini one dimensional plots. 



```{r}

dfa %>% 
  ggplot(aes(x = pct_partic_women, y = sport))+
  geom_violin()
```

